Testing Logistic Regression Models (Supplement to: Environmental DNA Metabarcoding Reveals Winners and Losers of Global Change in Coastal Waters)

### 2020-09-22 Scripts/Denoising.all.runs.Rmd
I think we good

 library (tidyverse)
 library (vegan)
 #library (MASS)
 library (proxy)
 library (reshape2)
 library (seqinr)
 library(patchwork)

@@

Hash.key %>% 
  distinct(Hash, .keep_all = T) -> Hash.key
```

@@

occ.results <- read_csv("../Input/Occ.fate.csv")

occ.results %>% 
  ggplot(aes(x = real)) +
  geom_histogram(fill = "blue") +
  labs(x = "",
       y = "Number of Hashes")  -> left.plot

occ.results %>% 
  left_join(ASV.nested %>% 
              select(Step2.tibble) %>% 
              unnest(cols = Step2.tibble) %>% 
              group_by(Hash) %>% 
              summarise (tot = sum(nReads))) %>% 
  ggplot(aes(x = cut_interval(real, n = 20))) +
  geom_col(aes(y = tot), fill = "red") +

@@

```{r actual filtering}
 occ.results %>% 
  filter(real > 0.8) %>% 
  pull (Hash) -> to.keep

@@

cleaned.tibble %>% 
  group_by(original_sample) %>% 
  summarise(nrep = n_distinct(sample)) %>% 
  #filter (nrep == 2) # 13
  filter (nrep == 1) # 0 
```
Ok, so there are 13 samples for which we only have 2 PCR replicates1.   We will get rid of those with only 1, as we can't estimate the PCR bias there.

@@

#names(all.distances.full)

summary(is.na(names(all.distances.full)))
```

@@

as.tibble(subset(melt(as.matrix(all.distances.full)))) -> all.distances.melted

summary(is.na(all.distances.melted$value))

@@

sapply(all.distances.to.plot, function(x) summary(is.na(x)))

@@

  ggplot (all.distances.to.plot , aes (fill = Distance.type, x = value)) +
  geom_histogram (position = "dodge", stat = 'density', alpha = 0.9) +
  facet_wrap( ~ Distance.type) +
  labs (x = "Pairwise dissimilarity", y = "density" ,
        Distance.type = "Distance")
ggsave("visual.anova.png", dpi = "retina")
```

@@

@@

Coverage.dataset %>% 
  ggplot()+
  geom_raster(aes(x= Date, y = Site, fill = cases))+
  geom_text(aes(x= Date, y = Site, label = cases))

>>>>

 library (tidyverse)
 library (vegan)
 #library (MASS)
 library (proxy)
 library (reshape2)
 library (seqinr)
 library (patchwork)

@@

Hash.key %>% 
  distinct(Hash, .keep_all = T) -> Hash.key

# Write the Hash key from start

Hash.key %>% 
  write_csv(here("Output", "starting.hash.key.csv"))

@@

occ.results <- read_csv("../Input/Occ.fate.2020.csv")

occ.results %>% 
  ggplot(aes(x = max)) +
  geom_histogram(fill = "blue") +
  labs(x = "",
       y = "Number of Hashes")  -> left.plot

occ.results %>% 
  left_join(ASV.nested %>% 
              select(Step2.tibble) %>% 
              unnest(cols = Step2.tibble) %>% 
              group_by(Hash) %>% 
              summarise (tot = sum(nReads))) %>% 
  ggplot(aes(x = cut_interval(max, n = 20))) +
  geom_col(aes(y = tot), fill = "red", position = "stack") +

@@

```{r actual filtering}
 occ.results %>% 
  filter(max > 0.8) %>% 
  pull (Hash) -> to.keep

@@

cleaned.tibble %>% 
  group_by(original_sample) %>% 
  summarise(nrep = n_distinct(sample)) %>% 
  #filter (nrep == 2) # 13
  filter (nrep == 1) # 0 
```
Ok, so there are 13 samples for which we only have 2 PCR replicates1.   We will get rid of those with only 1, as we can't estimate the PCR bias there. THis is

@@

# Do all samples have a name?
summary(is.na(names(all.distances.full))) # Yes they do
```

@@

as_tibble(subset(melt(as.matrix(all.distances.full)))) -> all.distances.melted

# Any mjor screw ups
summary(is.na(all.distances.melted$value))

@@

sapply(all.distances.to.plot, function(x) summary(is.na(x))) # good boi

@@

  ggplot (all.distances.to.plot ) +
  geom_histogram (aes (fill = Distance.type, x = value, after_stat(ndensity)), position = "dodge",  alpha = 0.9, bins = 50) +
  facet_wrap( ~ Distance.type) +
  labs (x = "Pairwise dissimilarity", y = "density" ,
        Distance.type = "Distance") +
    guides (fill = "none")
ggsave("visual.anova.png", dpi = "retina")
```

So our the distribution of dissimilarities is as we expected : lowest in technical replicates, then biological replicates and higher across our study system. Now let's see if there are any technical replicates that should be discarded due to their higher dissimilarity. We will calculate the distance from each PCR replicate to their group centroid, fit those distances to a normal distribution and discard values that are too high

@@

hist(all_distances)
```

@@

Coverage.dataset %>% 
  ggplot()+
  geom_raster(aes(x= Date, y = Site, fill = cases))+
  geom_text(aes(x= Date, y = Site, label = cases)) +
  facet_wrap(~Area, ncol = 1, scales = "free_y")

### 2020-09-22 Scripts/Manuscript.Main.Analysis.Rmd
I think we good

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflict_prefer("area", "patchwork")
#functions for future Salish Sea
models <- readRDS(here("Input","finalModel_20200807.RDS"))

@@

taxa.to.follow %>% 
  ggplot(aes(x = average)) + 
  geom_histogram() +
  facet_grid(Area ~ Comparison)
```

>>>>

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflict_prefer("area", "patchwork")
#functions for future Salish Sea
models <- readRDS(here("Input","bestModelFit.rds"))

@@

taxa.to.follow %>% 
  mutate(Comparison = case_when(Comparison %in% c("1_2", "2_1") ~ "1_2",
                                Comparison %in% c("1_3", "3_1") ~ "1_3",
                                Comparison %in% c("3_2", "2_3") ~ "2_3")) %>% 
  ggplot(aes(x = average)) + 
  geom_histogram() +
  facet_grid(Area ~ Comparison)
```

### 2020-09-22 Scripts/Prepare.data.for.stan.Rmd
I think we good

```{r}
library(tidyverse)

plankton <- read_csv(here("Input","Combined_Biol_Env_Plankton.csv"))
phylum.info <- read_csv(here("Input","higher_taxonomy.csv")) %>% 
  unite(family, genus, species, sep = "|", col = "taxa") %>% 
  select(phylum, taxa)

>>>>

```{r}
library(tidyverse)
old<- read_csv(here("Input","data.for.stan.csv"))
plankton <- read_csv(here("Input","Combined_Biol_Env_Plankton.csv"))
phylum.info <- read_csv(here("Input","higher_taxonomy.csv")) %>% 
  unite(family, genus, species, sep = "|", col = "taxa") %>% 
  select(phylum, taxa)

### 2020-09-22 Scripts/Rjags.tunning.Rmd
I think we good

We can modify the matrix by subsetting the data and doing the detection of an ASV in a month or Site - and later keep all hashes that were true (prob of Occ > 0.8) for a Site or Month.

We can modify the initial parameters by hard-coding a maximum value for the false positive rate; or by picking it up from a left-skewed beta distribution
```{r}
library(tidyverse)
library(vegan)
library(rjags)
library(proxy)
library(here)
library(unmarked)
library(jagsUI)

@@

@@

@@

all.hashes <- read_rds(here("Input", "Occupancy.nosite.rds"))

all.hashes ->test

>>>>

We can modify the matrix by subsetting the data and doing the detection of an ASV in a month or Site - and later keep all hashes that were true (prob of Occ > 0.8).

```{r}
library(tidyverse)
library(vegan)
library(rjags)
library(proxy)
library(here)
library(unmarked)
library(jagsUI)

@@

all.combos.10chain <- read_rds(here("Input", "10chain.rds"))

@@

row.df %>% 
  left_join(unique.by.site, by = c("Hash" = "first", "Site")) %>% 
  ungroup() %>% 
  select(-Hash, -Site) %>% 
  right_join(Pattern.of.presence) %>% 
  group_by(Hash) %>% tally(name = "max", sort = T)

@@

all.hashes <- read_rds(here("Input", "Occupancy.nosite.rds"))
```

```{r}
all.hashes %>% unnest(cols = c(data, starts_with("model")), names_repair = "unique") -> row1.df

row1.df %>% ungroup() %>% rowwise( Hash) %>% summarise(mean = mean(c_across(where(is.numeric))), 
                                                       median =median(c_across(where(is.numeric))),
                                                       max  = max(c_across(where(is.numeric)))) %>%
  left_join(Pattern.of.presence2, by = "Hash") %>% # this adds the output to the key
  ungroup %>% select(-Hash) %>% 
  left_join(Pattern.of.presence2) %>% write_csv(here("Input", "Occ.fate.2020.csv")) # this adds the output to the Hashes
```

### 2020-09-22 Scripts/Supplement_ModelTesting.Rmd
I think we good

```{r, echo = F, include = F}
load("bestModelFit.Rdata") #if loading from existing data
stan.input <- data.for.stan %>%
      group_by(Area) %>%
      mutate(Temperature = vegan::decostand(Temperature, method = "standardize"),
             pH = vegan::decostand(pH, method = "standardize"),
             Salinity = vegan::decostand(Salinity, method = "standardize")) %>%
      group_by(taxa) %>%
      mutate(Noccur = sum(binary)) %>%
      filter(Noccur > 7 & Noccur < 69) %>%  #must be present or absent in at least 10% of 76 samples in dataset
      dplyr::select(-c(sample, phylum, DIC, Noccur)) %>%
      ungroup() %>%
      unite(Area, taxa, col = "AreaTaxa", remove = F) %>%
      mutate(sampleIndex = 1:nrow(.),
             taxonIndex = match(taxa, unique(taxa)),
             areaIndex = match(Area, unique(Area)),
             seasonIndex = match(Season, unique(Season)),
             taxonAreaIndex = match(AreaTaxa, unique(AreaTaxa))) %>%   #unique combinations of taxon and area
      dplyr::select(-AreaTaxa)

#uncomment if running de novo; otherwise, load data from file, above
# finalModel <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa) + (1 | taxonAreaIndex),
#                              data = stan.input,
#                              family = "binomial",
#                              prior = t_prior,
#                              prior_intercept = t_prior,
#                              chains = 4,
#                              adapt_delta = 0.99,
#                              refresh = 0)
#optionally save
#save(finalModel, file = "bestModelFit.Rdata")
```

>>>>

```{r, echo = F, include = F}
load("bestModelFit.Rdata") #if loading from existing data
stan.input <- data.for.stan %>%
      group_by(Area) %>%
      mutate(Temperature = vegan::decostand(Temperature, method = "standardize"),
             pH = vegan::decostand(pH, method = "standardize"),
             Salinity = vegan::decostand(Salinity, method = "standardize")) %>%
      group_by(taxa) %>%
      mutate(Noccur = sum(binary)) %>%
      filter(Noccur > 7 & Noccur < 69) %>%  #must be present or absent in at least 10% of 76 samples in dataset
      dplyr::select(-c(sample, phylum, DIC, Noccur)) %>%
      ungroup() %>%
      unite(Area, taxa, col = "AreaTaxa", remove = F) %>%
      mutate(sampleIndex = 1:nrow(.),
             taxonIndex = match(taxa, unique(taxa)),
             areaIndex = match(Area, unique(Area)),
             #seasonIndex = match(Season, unique(Season)),
             taxonAreaIndex = match(AreaTaxa, unique(AreaTaxa))) %>%   #unique combinations of taxon and area
      dplyr::select(-AreaTaxa)

#uncomment if running de novo; otherwise, load data from file, above
finalModel <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa) + (1 | taxonAreaIndex),
                             data = stan.input,
                             family = "binomial",
                             prior = t_prior,
                             prior_intercept = t_prior,
                             chains = 4,
                             adapt_delta = 0.99,
                             refresh = 0)
# optionally save
write_rds(finalModel,path = here("Input","bestModelFit.rds"))
```
